---
title: "Be-Healthy"
author: "Pooja Rathee"
output:
  pdf_document: default
date: '2022-04-08'
---
## First we install required packages and go through it
```{r}
library(tidyverse)
```

## Read csv file
```{r}
daily_activity <- read_csv("dailyActivity_merged.csv")
sleep_day <- read.csv("sleepDay_merged.csv")
```

## Take a look at the daily activity data
```{r}
head(daily_activity)
```

# Identify all the columns in daily_activity
```{r}
colnames(daily_activity)
```
# Take a look at the sleep data
```{r}
head(sleep_day)
```

# Idenify all the columns in sleep day
```{r}
colnames(sleep_day)
```
# Understanding some summary statistics
```{r distinct users}
n_distinct(daily_activity$Id)
n_distinct(sleep_day$Id)

```
# Calculate the observations 
```{r observations}
nrow(daily_activity)
nrow(sleep_day)
```
# summary statistics of daily activity
```{r}
daily_activity %>%  
  select(TotalSteps,
         TotalDistance,
         SedentaryMinutes) %>%
  summary()
```
# For the sleep dataframe
```{r}
sleep_day %>%  
  select(TotalSleepRecords,
  TotalMinutesAsleep,
  TotalTimeInBed) %>%
  summary()
```
```{r}
ggplot(data=daily_activity, aes(x=TotalSteps, y=SedentaryMinutes)) + geom_point()+
  labs(title ="Relationship between total steps and Sedentary Minutes")
```
# plotting the graph for sleep_day
```{r}
ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + geom_point() +
  labs(title = "Relationship between total minutes asleep and total time in bed")
```
What's the relationship between minutes asleep and time in bed? You might expect it to be almost completely linear - are there any unexpected trends?

```{r}
ggplot(data=sleep_day, aes(x=TotalMinutesAsleep, y=TotalTimeInBed)) + geom_point()
```

What could these trends tell you about how to help market this product? Or areas where you might want to explore further?

## Merging these two datasets together

```{r}
combined_data <- merge(sleep_day, daily_activity, by="Id")
```

# How many participants are there in data
```{r}
n_distinct(combined_data$Id)
```


# There were more participant Ids in the daily activity dataset that have been filtered out using merge.


# Example change for git
