---
title: "Be-Healthy"
author: "Pooja Rathee"
output:
  pdf_document: default
date: '2022-04-08'
---
## First we install required packages and go through it
```{r}
library(tidyverse)
library(ggplot2)

```

## Read csv file
```{r}
daily_calories <- read_csv("data/dailyCalories_merged.csv")
weight_log <- read.csv("data/weightLogInfo_merged.csv")
```

## Take a look at the daily calories data
```{r}
head(daily_calories)
```

# Identify all the columns in daily_activity
```{r}
colnames(daily_calories)
```
# Take a look at the weight_log
```{r}
head(weight_log)
```

# Idenify all the columns in weight log
```{r}
colnames(weight_log)
```
# Understanding some summary statistics
```{r distinct users}
n_distinct(daily_calories$Id)
n_distinct(weight_log$Id)

```
# Calculate the observations 
```{r observations}
nrow(daily_calories)
nrow(weight_log)
```
# summary statistics of daily calories
```{r}
daily_calories %>%  
  select(Id,
         ActivityDay,
         Calories) %>%
  summary()
```
# For the weight_log dataframe
```{r}
weight_log %>%  
  select(Id,
  Date,
  WeightKg, BMI,Fat) %>%
  summary()
```
```{r}
ggplot(data=daily_calories, aes(x=ActivityDay, y=Calories, fill=Id)) + geom_point()+
  labs(title ="Relationship between ActivityDay and Calories Intake") 
```
# plotting the graph for weight_log
```{r}
ggplot(data=weight_log, aes(x=Date, y=BMI, fill=Id)) + geom_point() +
  labs(title = "Relationship between Date and BMI rate as per Id's")
```

What could these trends tell you about how to help market this product? Or areas where you might want to explore further?

## Merging these two datasets together
How Calories intake and BMI rate influences health?

```{r}
combined_data <- merge(weight_log, daily_calories, by="Id")
ggplot(data=combined_data,aes(x=Calories, y=BMI)) + 
  geom_line()
```

# How many participants are there in data
```{r}
ggplot(data=combined_data,aes(x=Id, y=BMI)) + 
  geom_line()
n_distinct(combined_data$Id)
```


# There were more participant Ids in the daily calories dataset that have been filtered out using merge.
