---
title: "Be-Healthy"
author: "Pooja Rathee"
output:
  pdf_document: default
date: '2022-04-19'
---

## First we install required packages and go through them

```{r}
library(tidyverse)
library(ggplot2) 
library(purrr)
library(readr)
library(scales)
library(reshape2)


```

#Importing the required data

```{r}

Activity <- read.csv("data/dailyActivity_merged.csv")
Calories<- read.csv("data/dailyCalories_merged.csv")
Intensities <- read.csv("data/dailyIntensities_merged.csv")
Steps <- read.csv("data/dailySteps_merged.csv")
sleepDay <- read.csv("data/sleepDay_merged.csv")
weight <- read.csv("data/weightLogInfo_merged.csv")

```

## A quick look in the following columns

```{r}
colnames(Activity)
colnames(Calories)
colnames(Intensities)
colnames(Steps)
colnames(sleepDay)
colnames(weight)

```

# Merging activity and sleep data

```{r}
merge_1 <- merge(Activity,Calories, by = c("Id","Calories"))
merge_2 <- merge(Intensities,Intensities, by = c("Id","ActivityDay","SedentaryMinutes", "LightlyActiveMinutes","FairlyActiveMinutes","VeryActiveMinutes", "SedentaryActiveDistance", "LightActiveDistance", "ModeratelyActiveDistance", "VeryActiveDistance"))

merge_daily <- merge(merge_1, merge_2, by = c("Id","ActivityDay","SedentaryMinutes", "LightlyActiveMinutes","FairlyActiveMinutes","VeryActiveMinutes", "SedentaryActiveDistance", "LightActiveDistance", "ModeratelyActiveDistance", "VeryActiveDistance")) %>%
select(-ActivityDay) %>% rename(Date = ActivityDate)

daily_data <- merge(merge_daily, sleepDay, by = "Id",all=TRUE) %>% drop_na() %>% select(-SleepDay, -TrackerDistance)


```

## Take a look on the summary

```{r}
summary(daily_data)

```

# Categorize the data on the basis of active minutes

```{r}
data_by_usertype <- daily_data %>%
summarise(
user_type = factor(case_when(
    SedentaryMinutes > mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Sedentary",
    SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes > mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Lightly Active",
    SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes > mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Fairly Active",
    SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes > mean(VeryActiveMinutes) ~ "Very Active",
),levels=c("Sedentary", "Lightly Active", "Fairly Active", "Very Active")), Calories, .group=Id) %>% drop_na()

```

# Visualize both the user type distribution and the calories burned for every user type:

```{r}
data_by_usertype %>%
group_by(user_type) %>%
summarise(total = n()) %>%
mutate(totals = sum(total)) %>%
group_by(user_type) %>%
summarise(total_percent = total / totals) %>%
ggplot(aes(user_type,y=total_percent, fill=user_type)) +
    geom_col()+
    scale_y_continuous(labels = scales::percent) +
    theme(legend.position="none") +
    labs(title="User type distribution", x=NULL) +
    theme(legend.position="none", text = element_text(size = 10),plot.title = element_text(hjust = 0.5))

ggplot(data_by_usertype, aes(user_type, Calories, fill=user_type)) +
    geom_boxplot() +
    theme(legend.position="none") +
    labs(title="Calories burned by User type", x=NULL) +
    theme(legend.position="none", text = element_text(size = 10),plot.title = element_text(hjust = 0.5))

```
The users are mostly sedentary or Lightly active users but it's very interesting to see that even though they are the biggest category the Fairly active and most importantly, the Very active are the ones with more calories burned. 
# To check the relation between Distance/Steps and Calories burned by plotting them:

```{r}
daily_data %>%
summarise(
distance = factor(case_when(
    TotalDistance < 4.5 ~ "< 4.5 mi",
    TotalDistance >= 4.5 & TotalDistance <= 7 ~ "4.5 > & < 7 mi",
    TotalDistance > 7 ~ "> 7 mi",
),levels = c("> 7 mi","4.5 > & < 7 mi","< 4.5 mi")),
steps = factor(case_when(
    TotalSteps < 6000 ~ "< 6k steps",
    TotalSteps >= 6000 & TotalSteps <= 10000 ~ "6k > & < 10k Steps",
    TotalSteps > 10000 ~ "> 10k Steps",
),levels = c("> 10k Steps","6k > & < 10k Steps","< 6k steps")),
Calories) %>%
ggplot(aes(steps,Calories,fill=steps)) +
    geom_boxplot() +
    facet_wrap(~distance)+
    labs(title="Calories burned by Steps and Distance",x=NULL) +
    theme(legend.position="none", text = element_text(size = 6),plot.title = element_text(hjust = 0.5))

```
The most calories burned are the "6k > & < 10k Steps" and "> 7 miles" which would indicate some kind of running activity that allows the user to traverse more distance with less steps.
# Understanding some summary statistics by visualization

```{r distinct users}
#Now let's focus on the sleep quality, for that I will now make categories for the sleeping time and I will make a new table with the sleeping categories percent for each individual user type:
sleepType_by_userType <- daily_data %>%
group_by(Id) %>%
summarise(user_type = factor(case_when(
    SedentaryMinutes > mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Sedentary",
    SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes > mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Lightly Active",
    SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes > mean(FairlyActiveMinutes) & VeryActiveMinutes < mean(VeryActiveMinutes) ~ "Fairly Active",
    SedentaryMinutes < mean(SedentaryMinutes) & LightlyActiveMinutes < mean(LightlyActiveMinutes) & FairlyActiveMinutes < mean(FairlyActiveMinutes) & VeryActiveMinutes > mean(VeryActiveMinutes) ~ "Very Active",
),levels=c("Sedentary", "Lightly Active", "Fairly Active", "Very Active")),
sleep_type = factor(case_when(
    mean(TotalMinutesAsleep) < 360 ~ "Bad Sleep",
    mean(TotalMinutesAsleep) > 360 & mean(TotalMinutesAsleep) <= 480 ~ "Normal Sleep",
    mean(TotalMinutesAsleep) > 480 ~ "Over Sleep"),levels=c("Bad Sleep", "Normal Sleep", "Over Sleep")), total_sleep = sum(TotalMinutesAsleep) ,.groups="drop"
) %>%
drop_na() %>%
group_by(user_type) %>%
summarise(bad_sleepers = sum(sleep_type == "Bad Sleep"), normal_sleepers = sum(sleep_type == "Normal Sleep"),over_sleepers = sum(sleep_type == "Over Sleep"),total=n(),.groups="drop") %>%
group_by(user_type) %>%
summarise(
    bad_sleepers = bad_sleepers / total, 
    normal_sleepers = normal_sleepers / total, 
    over_sleepers = over_sleepers / total,
    .groups="drop"

)


```

#Now we can plot the data for each user type:

```{r}
sleepType_by_userType_melted<- melt(sleepType_by_userType, id.vars = "user_type")

ggplot(sleepType_by_userType_melted, aes(user_type, value, fill = variable)) +
geom_bar(position = "dodge", stat = "identity") +
scale_y_continuous(labels = scales::percent) +
labs(x=NULL, fill="Sleep type") + 
theme(legend.position="bottom",text = element_text(size = 10),plot.title = element_text(hjust = 0.5))
```
It shows very clearly the relation between the activity level and sleep quality as in the sedentary users we find the largest percentage of bad sleepers and with some activity (even very little activity) we see a great increase of normal sleepers.


